Attract and kill: Spinosad containing spheres to control onion maggot (\textit{Delia antiqua})

### 2019-09-19 manuscript/pest-management-science/main.tex
More adjustments based on author feedback.

To enhance attractiveness, longevity, and performance of spinosad containing spheres, a new formulation of spinosad containing sphere was developed with 0.5\% spinosad, 5.0\% ammonium carbonate and 5.0\% casein hydrolysate. The efficacy of this new formulation was compared with spheres containing just 0.5\% spinosad and no spinosad controls in field cage trials. Treatments were deployed in a randomized controlled block design with four replications and each individual plot consisted of four rows on onion each 12 ft long. All onion seeds including those in the untreated control were treated with fungicides: thiram (Thiram 42 S {@}188 mg a.i./ 100 g seed) and tebuconazole (Raxil 2.6 F @ 250 mg a.i. / 100 g seed). Seeds were planted in early May and each plot was caged using PVC pipe and screen mesh. One hundred fifty \textit{D. antiqua} flies with a 50:50 male to female ratio were released into each cage on the second of June and again nine days later on June 11th. As described above, initial and final stand counts were taken and plant damage due to \textit{D. antiqua} larvae evaluated on a weekly basis.

@@

Deployment of the attract and kill solution of spinosad containing spheres did not substantially affect end of season onion loss as a result of all mortality factors (Figure \ref{fig:figure4}A), but did influence cumulative damage as a result of \textit{D. antiqua} larval feeding (Figure \ref{fig:figure4}B).  Treatment solution (Insecticide, No Insectide, Spinosad Sphere, Sticky Card) significantly (F = 5.8, df = 3, p = 0.005) explained approximately 38\% of the observed variation in end of season onion stand loss as a result of all mortality factors ($R^2_{adj}$ = 0.38, p = 0.005, Figure \ref{fig:figure4}A).  The presence of Delia Lure did not significantly (p \textgreater 0.05) affect end of season onion stand loss.  Insecticide treatment worked best at preventing end of season stand loss (Figure \ref{fig:figure4}A).  The no insecticide treatment was highly variable and not significantly (p \textgreater 0.05)  different from insecticide controls.  Spinosad and sticky card treatments were not significantly different from no insecticide treatments, but were significantly worse than insecticide controls.

Cumulative damage as a result of \textit{D. antiqua} larval feeding was significantly explained by treatment solution ($\chi^2$ = 42.2, df = 3, p \textless 0.001), the presence of Delia Lure ($\chi^2$ = 10.0, df = 1, p = 0.001), and their interaction ($\chi^2$ = 21.4, df = 1 p \textless 0.001).  Cumulative damage due to feeding by \textit{D. antiqua} larvae over the course of the field season ranged from 0 in some control plots to 82 dead \textit{D. antiqua} plants in a few plots. This amounts to approximately 20-30\% losses of onion directly attributable feeding by \textit{D. antiqua} larvae.  All treatments were significantly worse than insecticide controls which had very low levels of \textit{D. antiqua} damage (Figure \ref{fig:figure4}B).  Pairing Delia Lure with spinosad containing spheres significantly reduced damage by \textit{D. antiqua} larvae by 72$\pm$10\% (p \textless 0.001) compared to spinosad spheres alone.  Pairing Delia Lure with sticky cards had the opposite effect (Figure \ref{fig:figure4}B).

@@

Treatment significantly explained observed \textit{D. antiqua} damage in field cage trials with the enhanced formulation ($\chi^2$ = 23.9, df = 2, p \textless 0.001).  The enhanced formulation (spinosad +) containing 0.5\% spinosad, 5.0\% ammonium carbonate and 5.0\% casein hydrolysate did not perform significantly different than spinosad-only treatments with 0.5\% spinosad containing spheres.  Both treatments containing spinosad performed significantly better than the no insecticide controls (p \textless 0.05).

@@

\begin{figure}[bt]
\centering
\includegraphics[width = 8cm]{figures/final-figures/figure-2.pdf}
\caption{Adult onion maggot (\textit{D. antiqua} mortality in the presence of spinosad treated spheres.  (A) Raw data of effects of exposure time and spinosad rate on mortality of adult \textit{D. antiqua} flies.  Light points indicate individual observations.  Solid points and error bars denote mean and 95\% confidence intervals respectively.  (B) Main effects of spinosad rate on adult \textit{D. antiqua} mortality.  Points and error bars denote mean effect and 95\% confidence intervals respectively.  Different letters denote significant differences (p \textless 0.05).  (C) Main effect of exposure time on adult \textit{D. antiqua} mortality.  Line and shaded area denote effect and 95\% confidence region respectively.  (D) Photographs of spinosad treated spheres at increasing exposure time in field conditions.  Spheres were placed in the field on the 15th of May.  }
\label{fig:figure2}
\end{figure}

\begin{figure}[bt]
\centering
\includegraphics[width = 8cm]{figures/final-figures/figure-3.pdf}
\caption{Adult onion maggot fly mortality following introduction to cages containing spinosad treated spheres.  Spheres were treated with differing rates of spinosad (rows) and exposed to field conditions for increasing amounts of time (column).  Lines and shaded regions indicate fitted smoothed (LOESS) mortality rates and 95\% confidence regions respectivel.  }
\label{fig:figure3}
\end{figure}

\begin{figure}[bt]
\centering
\includegraphics[width = 8cm]{figures/final-figures/figure-4.pdf}
\caption{Field damage of onions.  (A) End of season stand loss of onions (number of dead onion plants) as a result of all mortality factors.  (B) Cumulative damage (number of dead onion plants) as a result of \textit{D. antiqua} larval feeding.  Light points indicate observed values.  Solid points and errorbars denote mean effects and 95\% confidence intevals respectively.  Different letters denote significant differences (p \textless 0.05).   }
\label{fig:figure4}
\end{figure}

\begin{figure}[bt]
\centering
\includegraphics[width = 8cm]{figures/final-figures/figure-5.pdf}
\caption{Field damage of onion in enclosed cages with differing spinosad sphere formulations.  Cumulative damage is number of dead plants as a result of \textit{D. antiqua} damage.  Points and error bars denote mean and 95\% confidence intervals respectively.  Different letters denote significant differences (p \textless 0.05).  }
\label{fig:figure5}
\end{figure}

>>>>

To enhance attractiveness, longevity, and performance of spinosad containing spheres, a new formulation of spinosad containing sphere was developed with 0.5\% spinosad, 5.0\% ammonium carbonate and 5.0\% casein hydrolysate. The efficacy of this new formulation was compared with spheres containing just 0.5\% spinosad and no spinosad controls in field cage trials. Treatments were deployed in a randomized controlled block design with four replications and each individual plot consisted of four rows of onion each 3.7m (12ft)  long. All onion seeds including those in the untreated control were treated with fungicides: thiram (Thiram 42 S {@}188 mg a.i./ 100 g seed) and tebuconazole (Raxil 2.6 F @ 250 mg a.i. / 100 g seed). Seeds were planted in early May and each plot was caged (3.7m x 3.7m x 2m) using PVC pipe and screen mesh. One hundred fifty \textit{D. antiqua} flies with a 50:50 male to female ratio were released into each cage on the second of June and again nine days later on June 11th. As described above, initial and final stand counts were taken and plant damage due to \textit{D. antiqua} larvae evaluated on a weekly basis.

@@

Deployment of the attract and kill solution of spinosad containing spheres did influence cumulative onion damage in the field as a result of \textit{D. antiqua} larval feeding (Figure \ref{fig:figure4}A).  Cumulative damage as a result of \textit{D. antiqua} larval feeding was significantly explained by treatment solution ($\chi^2$ = 42.2, df = 3, p \textless 0.001), the presence of Delia Lure ($\chi^2$ = 10.0, df = 1, p = 0.001), and their interaction ($\chi^2$ = 21.4, df = 1 p \textless 0.001).  Cumulative damage due to feeding by \textit{D. antiqua} larvae over the course of the field season ranged from 0 in some control plots to 82 dead \textit{D. antiqua} plants in a few plots. This amounts to approximately 20-30\% losses of onion directly attributable feeding by \textit{D. antiqua} larvae.  All treatments were significantly worse than insecticide controls which had very low levels of \textit{D. antiqua} damage (Figure \ref{fig:figure4}A).  Pairing Delia Lure with spinosad containing spheres significantly reduced damage by \textit{D. antiqua} larvae by 72$\pm$10\% (p \textless 0.001) compared to spinosad spheres alone.  Pairing Delia Lure with sticky cards had the opposite effect (Figure \ref{fig:figure4}A).

@@

Treatment significantly explained observed \textit{D. antiqua} damage in field cage trials with the enhanced formulation ($\chi^2$ = 23.9, df = 2, p \textless 0.001).  The enhanced formulation (spinosad +) containing 0.5\% spinosad, 5.0\% ammonium carbonate and 5.0\% casein hydrolysate did not perform significantly different than spinosad-only treatments with 0.5\% spinosad containing spheres (Figure \ref{fig:figure4}B).  Both treatments containing spinosad performed significantly better than the no insecticide controls (p \textless 0.05, Figure \ref{fig:figure4}B).

@@

\begin{figure}[bt]
\centering
\includegraphics[width = 8cm]{figures/final-figures/figure-2.pdf}
\caption{Adult onion maggot (\textit{D. antiqua}) mortality in the presence of spinosad treated spheres.  (A) Raw data of effects of exposure time and spinosad rate on mortality of adult \textit{D. antiqua} flies.  Light points indicate individual observations.  Solid points and error bars denote mean and 95\% confidence intervals respectively.  (B) Main effects of spinosad rate on adult \textit{D. antiqua} mortality.  Points and error bars denote mean effect and 95\% confidence intervals respectively.  Different letters denote significant differences (p \textless 0.05).  (C) Main effect of exposure time on adult \textit{D. antiqua} mortality.  Line and shaded area denote effect and 95\% confidence region respectively.  (D) Photographs of spinosad treated spheres at increasing exposure time in field conditions.  Spheres were placed in the field on the 15th of May.  }
\label{fig:figure2}
\end{figure}

\begin{figure}[bt]
\centering
\includegraphics[width = 8cm]{figures/final-figures/figure-3.pdf}
\caption{Adult onion maggot (\textit{D. antiqua}) fly mortality following introduction to cages containing spinosad treated spheres.  Spheres were treated with differing rates of spinosad (rows) and exposed to field conditions for increasing amounts of time (column).  Lines and shaded regions indicate fitted smoothed (LOESS) mortality rates and 95\% confidence regions respectivel.  }
\label{fig:figure3}
\end{figure}

\begin{figure}[bt]
\centering
\includegraphics[width = 8cm]{figures/final-figures/figure-4.pdf}
\caption{Field damage of onions.  (A) Cumulative damage (number of dead onion plants) as a result of \textit{D. antiqua} larval feeding.  Solid points and errorbars denote mean effects and 95\% confidence intevals respectively.  Different letters denote significant differences (p \textless 0.05).  (B) Field damage of onion in enclosed cages with differing spinosad sphere formulations.  Cumulative damage is number of dead plants as a result of \textit{D. antiqua} damage.  Points and error bars denote mean and 95\% confidence intervals respectively.  Different letters denote significant differences (p \textless 0.05). } 
\label{fig:figure4}
\end{figure}
